---
title: "ATE with Propensity Score Matching"
output:
  html_document:
    df_print: paged
---

# A. Propensity Score Matching

According to the paper,[^1] under this assumption:

[^1]: Dehejia, R. and Wahba, S. (1998) *Propensity score matching methods for non-experimental causal studies*. <doi:10.3386/w6829>.

$$
{Y(0), Y(1)} \perp\!\!\!\perp T | X
$$

we have:

$$
\hat{\tau} = E_{X_i|T_i =1}[E(Y_{i} | T_i = 1, X_i) - E(Y_{i}|T_i=1, X_i)]
$$

and

$$
\hat{\tau}_{|T_{i}=1} = E_{X_i|T_i =1}[E(Y_{i} | T_i = 1, X_i) - E(Y_{i}|T_i=1, X_i) | T_i=1]
$$

As covariates, X, can be high-dimensional, we use propensity score as a dimensional reduction tool.

**Using propensity score matching**

We define propensity score as

$$
p(X) = pr(Z = 1 | X),
$$

We have

$$
{Y(0), Y(1)} \perp\!\!\!\perp T | X = {Y(0), Y(1)} \perp\!\!\!\perp T | p(X)
$$

## i. PSID1

Following Dehejia and Wahba (1998), I will try to do 1-to-1 nearest propensity score matching and with replacement. The propensity score is estimated by a logistic model.

```{r}
source("scripts/pscore.R")
source("scripts/report.R")
source("scripts/matching.R")
```

#### **a. ATE:**

```{r}
y <- filtered_psid1$re78
z <- filtered_psid1$treat
x <- matrix(filtered_psid1$ps, nrow(filtered_psid1), 1)
```

```{r}
# Matching without bias adjustment
ate.psid1 <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = FALSE, 
      replace = TRUE, 
      ties = FALSE, 
      estimand="ATE",
      distance.tolerance = 0,
    )
```

```{r}
# Display result
ate.matched.psid1 <- filtered_psid1[c(ate.psid1$index.treated, ate.psid1$index.control), ]
summary(ate.psid1)
```

#### **b. ATT**

```{r}
# Matching without bias adjustment
att.psid1 <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = FALSE, 
      replace = TRUE, 
      ties = FALSE, 
      estimand="ATT",
      distance.tolerance = 0,
    )

att.matched.psid1 <- filtered_psid1[c(att.psid1$index.treated, att.psid1$index.control), ]
summary(att.psid1)
```

We examine matching quality by visualization.

```{r}
# Matching quality
plot_match(filtered_psid1$ps, ate.psid1$index.treated, ate.psid1$index.control)
```

```{r}
par(mfrow = c(2, 2), oma = c(0, 0, 2, 0))
  
# Matched Control
hist(ate.matched.psid1[ate.matched.psid1$treat == 0, ]$ps, 
     main = "Matched Control", 
     xlab = "Propensity Scores", 
     freq = FALSE, 
     breaks = 20)

# Original Control
hist(filtered_psid1[filtered_psid1$treat == 0, ]$ps, 
     main = "Original Control", 
     xlab = "Propensity Scores", 
     freq = FALSE, 
     breaks = 20)

# Matched Treated
hist(ate.matched.psid1[ate.matched.psid1$treat == 1, ]$ps, 
     main = "Matched Treated", 
     xlab = "Propensity Scores", 
     freq = FALSE, 
     breaks = 20)

# Original Treated
hist(filtered_psid1[filtered_psid1$treat == 1, ]$ps, 
     main = "Original Treated", 
     xlab = "Propensity Scores", 
     freq = FALSE, 
     breaks = 20)

# Add a single title for all plots
mtext("Propensity Score Distributions", outer = TRUE, cex = 1.5)

```

```{r}
MatchBalance(treat~ age + I(age^2) + education + I(education^2) + 
                married + nodegree + black + hispanic + 
                re74 + I(re74^2) + re75 + I(re75^2) + 
                u74 * black, data=filtered_psid1, match.out=ate.psid1)
```

As we can see, the treated units were well matched across all propensity scores. We can also reproduce the average treatment effect on the treated of Match() using these estimators[^2]:

[^2]: Abadie, Alberto and Guido Imbens. 2006. “Large Sample Properties of Matching Estimators for Average Treatment Effects.” *Econometrica* 74(1): 235-267.

$$
\hat{\tau}_M = \frac{1}{N} \sum_{W_i=1} \left( \hat{Y}_i(1) - \hat{Y}_i(0) \right) 
           = \frac{1}{N} \sum_{i=1}^N  (2W_i-1) \left(1 + \frac{K_M(i)}{M} \right) Y_i
$$

Where

$$
\hat{Y}_i(1) =
\begin{cases} 
\frac{1}{M} \sum_{j \in \mathcal{J}_M(i)} Y_j, & \text{if } W_i = 0, \\
Y_i, & \text{if } W_i = 1.
\end{cases}
$$

$$
\hat{Y}_i(0) =
\begin{cases} 
Y_i, & \text{if } W_i = 0, \\
\frac{1}{M} \sum_{j \in \mathcal{J}_M(i)} Y_j, & \text{if } W_i = 1.
\end{cases}
$$

and $K_M(i)$ the number of times unit i is used as a match given that M matches per unit are used. In our project, following Dehejia and Wahba (1998), I am going to use M = 1.

```{r}
# Extract Matched Data
matched.treated <- ate.matched.psid1$re78[ate.matched.psid1$treat==1]
matched.control <- ate.matched.psid1$re78[ate.matched.psid1$treat==0]

# Calculate Paired Differences
paired.diffs <- matched.treated - matched.control

# Calculate ATE
(ate <- mean(paired.diffs))
```

For bias adjustment estimator proposed by Abadie, Alberto and Guido Imbens[^3],

[^3]: Abadie, Alberto and Guido Imbens. 2006. “Large Sample Properties of Matching Estimators for Average Treatment Effects.” *Econometrica* 74(1): 235-267.

```{r}
# Matching with bias adjustment
ate.psid1.AI <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = TRUE, 
      replace = TRUE, 
      estimand="ATE",
      distance.tolerance = 0,
    )
summary(ate.psid1.AI)
```

```{r}
# Matching with bias adjustment
att.psid1.AI <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = TRUE, 
      replace = TRUE, 
      estimand="ATT",
      distance.tolerance = 0,
    )

summary(att.psid1.AI)
```

Let's apply it for other dataset.

### ii. CPS1

#### a. ATE

```{r}
# Perform matching
filtered_cps1 <- pscore_cps1$filter_data
filtered_cps1$ps <- pscore_cps1$pscore
y <- filtered_cps1$re78
z <- filtered_cps1$treat
x <- matrix(filtered_cps1$ps, nrow(filtered_cps1), 1)
```

```{r}
ate.cps1 <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = FALSE, 
      replace = TRUE, 
      ties = FALSE, 
      estimand = "ATE"
    )

summary(ate.cps1)
```

```{r}
# Visualize matched pairs
plot_match(filtered_cps1$ps, ate.cps1$index.treated, ate.cps1$index.control)
```

```{r}
# Perform matching for unbiased ATE estimator
ate.cps1.AI <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      replace = TRUE, 
      estimand="ATE"
    )

summary(ate.cps1.AI)
```

#### b. ATT

```{r}
# Perform matching for ATT biased estimator
att.cps1 <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      BiasAdjust = FALSE,
      ties = FALSE,
      replace = TRUE, 
      estimand="ATT"
    )

summary(att.cps1)
```

```{r}
# Perform matching for unbiased ATT estimator
att.cps1.AI <- Match(
      Y = y, 
      Tr = z, 
      X = x, 
      M = 1, 
      replace = TRUE, 
      estimand="ATT"
    )

summary(att.cps1.AI)
```

### iii. Other datasets

```{r}
cat("1. ATE WITHOUT Bias Adjustment:\n")
ate.psid2 <- ps_matching_analysis(psid2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75)

cat("2. ATE WITH Bias Adjustment:\n")
ate.psid2.AI <- ps_matching_analysis(psid2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, bias_adjust = TRUE)

cat("3. ATT WITHOUT Bias Adjustment:\n")
att.psid2 <- ps_matching_analysis(psid2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, estimand="ATT")

cat("4. ATT WITH Bias Adjustment:\n")
att.psid2.AI <- ps_matching_analysis(psid2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, bias_adjust = TRUE, estimand="ATT")
```

```{r}
cat("1.ATE WITHOUT Bias Adjustment:\n")
ate.psid3 <- ps_matching_analysis(psid3_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75)

cat("2. ATE Bias Adjustment:\n")
ate.psid3.AI <- ps_matching_analysis(psid3_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, TRUE)

cat("3. ATT Bias Adjustment:\n")
att.psid3 <- ps_matching_analysis(psid3_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, estimand="ATT")

cat("4. ATT Bias Adjustment:\n")
att.psid3.AI <- ps_matching_analysis(psid3_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75, TRUE)
```

```{r}
cat("1. ATE WITHOUT Bias Adjustment:\n")
ate.cps2 <- ps_matching_analysis(cps2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75 + education*re74 + I(age^3))

cat("2. ATE WITH Bias Adjustment:\n")
ate.cps2.AI <- ps_matching_analysis(cps2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75 + education*re74 + I(age^3), bias_adjust=TRUE)

cat("3. WITHOUT Bias Adjustment:\n")
att.cps2 <- ps_matching_analysis(cps2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75 + education*re74 + I(age^3), estimand="ATT")

cat("4. WITH Bias Adjustment:\n")
att.cps2.AI <- ps_matching_analysis(cps2_re74, formula=treat ~ age + I(age^2) + education + I(education^2) + 
    married + nodegree + black + hispanic + 
    re74 + re75 + u74 + u75 + education*re74 + I(age^3), bias_adjust=TRUE, estimand="ATT")
```

```{r}
cat("1. ATE WITHOUT Bias Adjustment:\n")
ate.cps3 <- ps_matching_analysis(cps3_re74, formula=treat ~ age + I(age^2) + I(age^3) + education + I(education^2) + married + nodegree + black + hispanic + re74 + re75 + u74 + u75 + education*re74)

cat("2. ATE WITH Bias Adjustment:\n")
ate.cps3.AI <- ps_matching_analysis(cps3_re74, formula=treat ~ age + I(age^2) + I(age^3) + education + I(education^2) + married + nodegree + black + hispanic + re74 + re75 + u74 + u75 + education*re74, TRUE, estimand="ATT")

cat("3. ATT WITHOUT Bias Adjustment:\n")
att.cps3 <- ps_matching_analysis(cps3_re74, formula=treat ~ age + I(age^2) + I(age^3) + education + I(education^2) + married + nodegree + black + hispanic + re74 + re75 + u74 + u75 + education*re74)

cat("4. ATT WITH Bias Adjustment:\n")
att.cps3.AI <- ps_matching_analysis(cps3_re74, formula=treat ~ age + I(age^2) + I(age^3) + education + I(education^2) + married + nodegree + black + hispanic + re74 + re75 + u74 + u75 + education*re74, TRUE, estimand="ATT")
```

## Summary

In summary, propensity score matching improves the results. The bias-adjusted estimators produced higher standard error as expected. However, compared to the findings of the original paper, my result is not as strong.

What I have learnt so far:

-   In randomization, we can estimate ATE by $\hat{\tau} = E[Y(1) - Y(0)] = E[Y(1)] - E[Y(0)] = E[Y(1)|Z=1] - E[Y(0)|Z=0]$ because $\{Y(0), Y(1)\}$ is independent of Z and the linearity of expectation. However in observational study, that assumption is violated, so using that estimator will yield biases. But under strong ignorability assumption, that is $\{Y(1), Y(0)\}$ is independent of Z\|X, $E[Y(1) - Y(0) | X] = E[Y(1)|Z=1, X] - E[Y(0)|Z=0, X] = \tau(X)$, which can reduce the bias. Then to estimate $\hat{\tau}$, we can average this over the marginal distribution of X, which is $\hat{\tau} = E[\tau(X)]$.

-   As X can have high dimensions, propensity score matching estimators (Rosenbaum and Rubin (1983)) are widely used to estimate treatment effects.[^4] When the relevant differences between any two units are captured in the observed covariates, matching methods can yield an unbiased estimate of the treatment impact.[^5] Therefore, propensity score matching aims to balance the covariate differences between treated and control groups.

-   As we can never observe Yi(1) and Yi(0) together, we can use matching to impute the missing potential outcomes.[^6]

-   Why are ATTs so different from ATEs?

    -   For ATE, for every unit, if it is a treated unit, we match them with M control units, if it is a control unit, we match them with M treated unit. Then we compute the estimator similarly to the experimental study.

    -   While for ATT, for we only match every treated unit with M control units, and compute the average of unit-level treated effect.

    -   Additionally, there are not many treated units, so not all matches for control units are likely to be good.

    -   Hence, if the two groups are very different, more matches possibly mean producing more inexact matches, potentially causing bias in the estimators.

-   How can I improve the result

    -   Assessing the quality of the resulting matched samples, and perhaps iterating with steps 1 and 2 until well-matched samples result.[^7]

    -   More analysis of the outcome.

[^4]: Abedie and Imbens

[^5]: Dehejia (1998)

[^6]: Abedie and Imbens (2006)

[^7]: Elizabeth A. Stuart, Matching Methods for Causal Inference: A Review and a Look Forward

### i. ATE

```{r}
ate.psid1.out <- list(est=ate.psid1$est, se=ate.psid1$se.standard)
ate.cps1.out <- list(est=ate.cps1$est, se=ate.cps1$se.standard)

a <- list(ate.psid1.out, ate.psid2, ate.psid3, ate.cps1.out, ate.cps2, ate.cps3)

row_names <- c("NSW", "PSID-1", "PSID-2", "PSID-3", "CPS-1", "CPS-2", "CPS-3")
col_names <- c("Unadjusted PSM ATE")

(ate.psm <- create_psm_result(a, row_names, col_names))
```

```{r}
ate.psid1.AI.o <- list(est=ate.psid1.AI$est, se=ate.psid1.AI$se.standard)
ate.cps1.AI.o <- list(est=ate.cps1.AI$est, se=ate.cps1.AI$se.standard)

a <- list(ate.psid1.AI.o, ate.psid2.AI, ate.psid3.AI, ate.cps1.AI.o, ate.cps2.AI, ate.cps3.AI)

row_names <- c("NSW", "PSID-1", "PSID-2", "PSID-3", "CPS-1", "CPS-2", "CPS-3")
col_names <- c("Bias-Adjusted PSM ATE")

(ate.psm.AI <- create_psm_result(a, row_names, col_names))
```

```{r}
ate.result_table <- cbind(effects_table, ate.psm, ate.psm.AI)
row.names(ate.result_table) <- row_names
ate.result_table
```

### ii. ATT

```{r}
att.psid1.out <- list(est=att.psid1$est, se=att.psid1$se.standard)
att.cps1.out <- list(est=att.cps1$est, se=att.cps1$se.standard)

a <- list(att.psid1.out, att.psid2, att.psid3, att.cps1.out, att.cps2, att.cps3)

row_names <- c("NSW", "PSID-1", "PSID-2", "PSID-3", "CPS-1", "CPS-2", "CPS-3")
col_names <- c("Unadjusted PSM ATT")

(att.psm <- create_psm_result(a, row_names, col_names))
```

```{r}
att.psid1.AI.o <- list(est=att.psid1.AI$est, se=att.psid1.AI$se.standard)
att.cps1.AI.o <- list(est=att.cps1.AI$est, se=att.cps1.AI$se.standard)

a <- list(att.psid1.AI.o, att.psid2.AI, att.psid3.AI, att.cps1.AI.o, att.cps2.AI, att.cps3.AI)

row_names <- c("NSW", "PSID-1", "PSID-2", "PSID-3", "CPS-1", "CPS-2", "CPS-3")
col_names <- c("Bias-Adjusted PSM ATT")

(att.psm.AI <- create_psm_result(a, row_names, col_names))
```

```{r}
att.result_table <- cbind(effects_table, att.psm, att.psm.AI)
row.names(att.result_table) <- row_names
att.result_table
```
