---
title: "Average Treatment Effect in the National Supported Work Demonstration"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

## Load and create dataset

I will use these datasets from Dehejia and Wahba (1998)[^1]:

[^1]: Dehejia and Wahba

-   **The National Supported Work (NSW) Demonstration**

    -   RE74: include 2 years of pre-intervention earnings

    -   RE75: include 1 years of pre-intervention earnings

-   **PSID1-3 and CSP1-3**

    -   PSID-1: All male household heads less than 55 years old who did not classify themselves as retired in 1975.

    -   PSID-2: Selects from PSID-1 all men who were not working when surveyed in the spring of 1976.

    -   PSID-3: Selects from PSID-2 all men who were not working in 1975.

    -   CPS-1: All CPS males less than 55 years of age.

    -   CPS-2: Selects from CPS-1 all males who were not working when surveyed in March 1976.

    -   CPS-3: Selects from CPS-2 all the unemployed males in 1976 whose income in 1975 was below the poverty level.

    -   PSID1-3 and CPS-1 are identical to those used in Lalonde. CPS 2-3 are similar to those used in Lalonde, but Lalonde’s original subset could not be re-created.

Data Source: <https://users.nber.org/~rdehejia/nswdata2.html>

Now, let's check if the data is similar to the one in the paper.

**a. NSW dataset**

```{r}
# Load NSW data
nsw_tr <- read.table("data/nsw_treated.txt", header = FALSE)
nsw_co <- read.table("data/nsw_control.txt", header = FALSE)

# Assign column names
col_names <- c("treat", "age", "education", "black", "hispanic", "married", "nodegree", "re75", "re78")

colnames(nsw_tr) <- col_names
colnames(nsw_co) <- col_names
```

```{r}
# Summary statistics for treated
paste("Number of treated instances:", nrow(nsw_tr))
sapply(nsw_tr, \(col) round(mean(col), 2))

# Summary statistics for control
paste("Number of control instances:", nrow(nsw_co))
sapply(nsw_co, \(col) round(mean(col), 2))
```

As expected for an experimental dataset, the covariates are balance between treated and control groups.

**b. Re74 subset** This dataset is also experimental but includes a new covariate which is income in 1974 (Dehejia and Wahba).

```{r}
# Load Re74 data
re74_tr <- read.table("data/nswre74_treated.txt", header = FALSE)
re74_co <- read.table("data/nswre74_control.txt", header = FALSE)

# Assign column names
col_names <- c("treat", "age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75", "re78")
colnames(re74_tr) <- c(col_names)
colnames(re74_co) <- c(col_names)

```

```{r}
# Summary statistics for treated
paste("Number of treated instances (Re74):", nrow(re74_tr))
sapply(re74_tr, \(col) round(mean(col, na.rm = TRUE), 2))

# Summary statistics for control
paste("Number of control instances (Re74):", nrow(re74_co))
sapply(re74_co, \(col) round(mean(col, na.rm = TRUE), 2))
```

**b. Comparison groups**

```{r}
# Load PSID and CPS datasets
psid1 <- read.table("data/psid_controls.txt", header = FALSE)
psid2 <- read.table("data/psid2_controls.txt", header = FALSE)
psid3 <- read.table("data/psid3_controls.txt", header = FALSE)

cps1 <- read.table("data/cps_controls.txt", header = FALSE)
cps2 <- read.table("data/cps2_controls.txt", header = FALSE)
cps3 <- read.table("data/cps3_controls.txt", header = FALSE)
```

```{r}
# Assign column names
list_of_dfs <- list(psid1, psid2, psid3, cps1, cps2, cps3)

list_of_dfs <- lapply(list_of_dfs, function(df) {
  colnames(df) <- col_names
  return(df)
})

# Reassign the modified data frames
psid1 <- list_of_dfs[[1]]
psid2 <- list_of_dfs[[2]]
psid3 <- list_of_dfs[[3]]
cps1 <- list_of_dfs[[4]]
cps2 <- list_of_dfs[[5]]
cps3 <- list_of_dfs[[6]]
```

```{r}
data_name <- c("PSID-1", "PSID-2", "PSID-3", "CPS-1", "CPS-2", "CPS-3")
for (i in 1:length(list_of_dfs)) {
  print(data_name[i])
  df = list_of_dfs[[i]]
  
  means <- sapply(df, \(col) round(mean(col, na.rm = TRUE), 2))
  print(means)
}
```

The covariates are very different from the experimental one.

## The NSW Treatment Earnings Less Comparison Group Earnings

In this section, I try to estimate the average treatment effect (ATE) using both the original and observational samples without controlling covariates[^2].

[^2]: Peng Ding

$$
\hat{\tau} = \hat{\overline{Y}}(1) - \hat{\overline{Y}}(0)
$$

### 1. NSW treated group + Observational data

```{r}
source("scripts/ate_f.R")
```

```{r}
# Prepare data
nsw <- rbind(nsw_co, nsw_tr)

psid1_re75 <- rbind(nsw_tr, subset(psid1, select = -c(re74)))
psid2_re75 <- rbind(nsw_tr, subset(psid2, select = -c(re74)))
psid3_re75 <- rbind(nsw_tr, subset(psid3, select = -c(re74)))

cps1_re75 <- rbind(nsw_tr, subset(cps1, select = -c(re74)))
cps2_re75 <- rbind(nsw_tr, subset(cps2, select = -c(re74)))
cps3_re75 <- rbind(nsw_tr, subset(cps3, select = -c(re74)))
```

```{r}
# Estimate treatment effects
effects <- list(
  NSW = ate.experimental(nsw)$est,
  
  PSID1 = ate.experimental(psid1_re75)$est,
  PSID2 = ate.experimental(psid2_re75)$est,
  PSID3 = ate.experimental(psid3_re75)$est,
  
  CPS1 = ate.experimental(cps1_re75)$est,
  CPS2 = ate.experimental(cps2_re75)$est,
  CPS3 = ate.experimental(cps3_re75)$est
)
```

```{r}
effects_table <- do.call(rbind, effects)
rownames(effects_table) <- names(effects)

effects_table
```

### 2. LDW treated group without RE74

```{r}
# Prepare data
re74 <- rbind(re74_co, re74_tr)

psid1_re74 <- rbind(re74_tr, psid1)
psid2_re74 <- rbind(re74_tr, psid2)
psid3_re74 <- rbind(re74_tr, psid3)

cps1_re74 <- rbind(re74_tr, cps1)
cps2_re74 <- rbind(re74_tr, cps2)
cps3_re74 <- rbind(re74_tr,cps3)
```

```{r}
effects2 <- list(
  NSW = ate.experimental(subset(re74, select = -c(re74)))$est,
  
  PSID1 = ate.experimental(subset(psid1_re74, select = -c(re74)))$est,
  PSID2 = ate.experimental(subset(psid2_re74, select = -c(re74)))$est,
  PSID3 = ate.experimental(subset(psid3_re74, select = -c(re74)))$est,
  
  CPS1 = ate.experimental(subset(cps1_re74, select = -c(re74)))$est,
  CPS2 = ate.experimental(subset(cps2_re74, select = -c(re74)))$est,
  CPS3 = ate.experimental(subset(cps3_re74, select = -c(re74)))$est
)

```

```{r}
effects_table2 <- do.call(rbind, effects2)
rownames(effects_table2) <- names(effects2)

effects_table2
```

### 3. LDW treated group with RE74

```{r}
effects <- list(
  NSW = ate.experimental(re74)$est,
  
  PSID1 = ate.experimental(psid1_re74)$est,
  PSID2 = ate.experimental(psid2_re74)$est,
  PSID3 = ate.experimental(psid3_re74)$est,
  
  CPS1 = ate.experimental(cps1_re74)$est,
  CPS2 = ate.experimental(cps2_re74)$est,
  CPS3 = ate.experimental(cps3_re74)$est
)

```

```{r}
effects_table <- do.call(rbind, effects)
rownames(effects_table) <- names(effects)

effects_table
```

## Summary

-   The treatment effect is higher in Panel B.

-   The simple difference in means is negative for PSID1-2 and CPS1-3.

-   Overall, the results of observational studies are far from the result of experimental study.
