---
title: "ATE with Inverse Propensity Score Weighting"
output: html_notebook
---

# B. Inverse Propensity Score Weighting

We have this inverse propensity score weighting estimator[^1].

[^1]: Ding, Peng. (2023). A First Course in Causal Inference. 10.48550/arXiv.2305.18793.

$$
\widehat{\tau}^{\text{ht}} = n^{-1} \sum_{i=1}^n \frac{Z_iY_i}{\hat{e}(X_i)} - n^{-1} \sum_{i=1}^n \frac{(1 - Z_i)Y_i}{1 - \hat{e}(X_i)}
$$

and

$$
\widehat{\tau}^{\text{ht}}_{|T_i=1} = n_1^{-1} \sum_{i=1}^n Z_iY_i - n_1^{-1} \sum_{i=1}^n \frac{\hat{e}(X_i)(1 - Z_i)Y_i}{1 - \hat{e}(X_i)}
$$

We will use the propensity score estimated by a single index model.

```{r}
source("scripts/pscore.R")
source("scripts/ipw.R")
source("scripts/report.R")
```

```{r}
# IPW ATE
ate.psid1.ipw <- ps_IPW(filtered_psid1_SI, estimand="ATE", norm=FALSE)
paste("IPW ATE: ", ate.psid1.ipw$est)
```

```{r}
# IPW ATE
att.psid1.ipw <- ps_IPW(filtered_psid1_SI, estimand="ATT", norm=FALSE)
paste("IPW ATT: ", att.psid1.ipw$est)
```

Let's check distribution of the weights.

```{r}
# Extract weights and treatment
weights <- ate.psid1.ipw$weights
treat <- filtered_psid1_SI$treat

# Split weights by treatment group
weights_control <- weights[treat == 0]
weights_treated <- weights[treat == 1]

# Define histogram breaks
breaks <- seq(min(weights), max(weights), length.out = 31)

# Plot histograms
hist(weights_control, breaks = breaks, col = rgb(0, 0, 1, 0.6),
     main = "Distribution of Weights by Treatment Group",
     xlab = "Weights", ylab = "Count", xlim = range(weights))
hist(weights_treated, breaks = breaks, col = rgb(1, 0, 0, 0.6),add = TRUE)

# Add legend
legend("topright", legend = c("Control", "Treated"),
       fill = c(rgb(0, 0, 1, 0.6), rgb(1, 0, 0, 0.6)))

```

Some weights are very large, which may violate the overlap condition.

```{r}
# Create a list of datasets after discarding controls
SI_datasets <- list(filtered_psid1_SI, filtered_cps1_SI)
LM_datasets <- list(filtered_psid1, filtered_cps1)
```

```{r}
# Compute inverse propensity score weighting ATE for main datasets
ate_result.SI <- lapply(SI_datasets, function(data) {
  ps_IPW(data, estimand = "ATE")
})

# Compute inverse propensity score weighting ATE for main datasets
ate_result.logistic <- lapply(LM_datasets, function(data) {
  ps_IPW(data, estimand = "ATE")
})
```

```{r}
# Display result
row_names <- c("NSW", "PSID-1", "CPS-1")
col_names <- c("IPW ATE")
(ate.ipw.results.SI <- create_psm_result(ate_result.SI, row_names, col_names))
(ate.ipw.results.logistic <- create_psm_result(ate_result.logistic, row_names, col_names))
```

```{r}
# Compute inverse propensity score weighting ATT for main datasets
att_result.SI <- lapply(SI_datasets, function(data) {
  ps_IPW(data, estimand = "ATT")
})

# Compute inverse propensity score weighting ATT for main datasets
att_result.logistic <- lapply(LM_datasets, function(data) {
  ps_IPW(data, estimand = "ATT")
})
```

```{r}
# Display result
col_names <- c("IPW ATT")
(att.ipw.results.SI <- create_psm_result(att_result.SI, row_names, col_names))
(att.ipw.results.logistic <- create_psm_result(att_result.logistic, row_names, col_names))
```

## Summary

It has been proved that

$$
\tau = E\{Y(1) - Y(0)\} = E \left\{ \frac{ZY}{e(X)} - \frac{(1-Z)Y}{1-e(X)} \right\}.
$$

Under strong ignorability and overlap condition. As the formulas can be very big if the propensity score is very close to 1 or 0, the weighting estimators blow up to infinity resulting in extremely unstable behaviors in finite samples.[^2] As a result, the more we trimmed the psid1 dataset, the better estimate we got.

[^2]: Peng Ding

-   Treated subjects with a very low propensity score can result in a very large weight. Similarly, a control subject with a propensity score close to one can result in a very large weight. Such weights can increase the variability of the estimated treatment effect.[^3] (Why?)

[^3]: P. C. AUSTIN AND E. A. STUART

#### 1. What impacts the IPW estimates -TODO

Most of the extremely large weights correspond to treated units with very small propensity score.

#### 2. IPW vs Matching -TODO
