---
title: "Quantile Treatment Effect"
output: html_notebook
---

# Quantile Treatment Effect

We will approach QTE and QTT in two ways.

The first method is to use the estimator proposed by Firpo (2007).[^1] The first step is to estimate the propensity score using a nonparametric method. In his paper, he employed HIR method, which used a logistic power series approximation. The second step is to minimize a weighted sum of check functions.

[^1]: Sergio Firpo. (2007). Efficient Semiparametric Estimation of Quantile Treatment Effects. *Econometrica*, *75*(1), 259–276. <http://www.jstor.org/stable/4123114>

The second method is to use IPW estimators for $F_0(z)$ and $F_1(z)$.[^2] From the estimated distribution of the potential outcomes, we can directly estimate the quantile.

[^2]: Donald, S. G., & Hsu, Y.-C. (2014). Estimation and inference for distribution functions and quantile functions in treatment effect models. *Journal of Econometrics, 178*(3), 383–397. <https://doi.org/10.1016/j.jeconom.2013.03.010>

## 1. Firpo Approach

### a. Estimate Propensity Score using Series Logit Estimator

```{r}
load("data/ps_lalonde.RData")
source("scripts/report.R")
library(qte)
```

```{r}
# Data for testing
data <- psid1_re74
covariates <- c("age", "education", "re74") 
```

```{r}
# Generate polynomial basis returning vector H_X(K)
# cov refers to numerical covariates that we want to have higher terms
generate_polynomial_basis <- function(data, degree, cov) {
  
  H_K <- cbind(data, data.frame(Intercept = 1))  # Add the constant term

  for (cov in covariates) {
    poly_terms <- as.data.frame(poly(data[[cov]], degree = degree, raw = TRUE)[, -1])
    colnames(poly_terms) <- paste0(cov, "_", seq_len(ncol(poly_terms)) + 1)
    H_K <- cbind(H_K, poly_terms) 
  }
  
  return(as.data.frame(H_K))
}

degree <- 2
data_H_K <- generate_polynomial_basis(data, degree, covariates)
data_H_K
```

```{r}
# Objective function to find the propensity scores
objective_func <- function(pi, H_K, treat) {
  N <- nrow(H_K)
  z <- rowSums(H_K * pi)  # H_K(x)' * pi
  p <- 1 / (1 + exp(-z))
  p <- pmax(pmin(p, 1 - 1e-15), 1e-15)  # Bound p
  log_lik <- (treat * log(p) + (1 - treat) * log(1 - p))
  return(-sum(log_lik)/N) 
}
```

```{r}
# Estimate pi using minimize the objective function
estimate_pi <- function(H_K, treat, pi.init = NULL) {
  
  result <- optim(
    par = pi.init,
    fn = objective_func,
    H_K = H_K,
    treat = treat,
    method = "BFGS" 
  )
  
  return(result$par)
}
```

```{r}
# Compute propensity scores
compute_propensity_scores <- function(H_K, pi_K) {
  z <- rowSums(H_K * pi_K)  # H_K(x)' * pi
  p <- 1 / (1 + exp(-z))
  return(p)
}
```

```{r}
# Full propensity score estimation pipeline
firpo_propensity_scores <- function(data) {
  
  cov <- setdiff(colnames(data), c("treat", "re78")) # exclude 'treat' and "re78"
  H_K <- data[cov]
  treat <- data[["treat"]]
  pi.init <- rep(0, length(cov))
  
  pi_K <- estimate_pi(H_K, treat, pi.init)
  
  # Compute propensity scores
  propensity_scores <- compute_propensity_scores(H_K, pi_K)
  
  return(list(coefficients = pi_K, ps = propensity_scores))
}

```

```{r}
# Firpo ps model for psid1 dataset - page 6 in the supplement 
cols_up_to_degree_2 <- c(
  "treat", "age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75", "u74", "u75", "Intercept", 
  "age_2", "education_2", "re74_2")

ps1_fibro <- firpo_propensity_scores(data_H_K)$ps
```

```{r}
min_treat_ps <- min(ps1_fibro[data_H_K$treat == 1])
max_treat_ps <- max(ps1_fibro[data_H_K$treat == 1])
  
filter_mask <- (ps1_fibro > min_treat_ps & ps1_fibro < max_treat_ps) | data_H_K$treat == 1
ps1_fibro_trim <- ps1_fibro[filter_mask]
summary(ps1_fibro_trim)
```

```{r}
# Create probability bins
bins <- cut(ps1_fibro, breaks = seq(0, 1, by = 0.1), right = FALSE)

# Count treated and control units in each bin
counts_treat_0 <- table(bins[data_H_K$treat == 0])
counts_treat_1 <- table(bins[data_H_K$treat == 1])

# Combine counts
counts_matrix <- rbind(counts_treat_0, counts_treat_1)

# Generate bin labels
group_labels <- paste(seq(0.1, 1, by = 0.1))

# Create bar plot
barplot(
  counts_matrix,
  beside = TRUE,
  col = c("white", "blue"),
  legend.text = c("Control", "Treated"),
  main = "Histogram of Estimated Propensity Score",
  xlab = "Probability Bin",
  ylab = "Frequency",
  names.arg = group_labels,
  ylim = c(0, 200)
)
```

The code is very wrong.

### b. Estimate Quantile Treatment Effect

Check function is defined as

$$
p_r(a) = a \cdot ( \tau - \mathbb{1}(a \leq 0))$$

```{r}
# Define the check function
# Source code - https://github.com/bcallaway11/BMisc/blob/master/R/BMisc.R
# https://rdrr.io/cran/qte/src/R/ciqte.R
# I implement it again for studying and trimming purpose
check_function <- function(a, tau) {
  return(a * (tau - (1 * (a <= 0))))
}
```

To find $\hat{q}_{1,r}$ , we minimize

$$ \frac{1}{N} \sum_{i=1}^N \frac{T_i}{\hat{p}(X_i)} \cdot(Y_i - q) \cdot (\tau - 1\{Y_i < q\})$$

```{r}
# Define the quantile objective function
q_objective_func <- function(q, y, tau, weights) {
  return(mean(weights * check_function(y - q, tau)))
}
```

```{r}
# Estimate the `tau` quantile 
estimate_quantile <- function(tau, y, weights) {
  optim(
    par = median(y),
    fn = q_objective_func,
    y = y,
    tau = tau,
    weights = weights,
    method = "Brent",
    lower = min(y),
    upper = max(y)
  )$par
}
```

We estimate QTE as[^3]

[^3]: **Firpo, S.** (2007). Supplement to "Efficient Semiparametric Estimation of Quantile Treatment Effects." *Econometrica*, 75(1), 259–276, p. 261.

$$
QTE = \hat{q}_{1, \tau} - \hat{q}_{0, \tau}
$$

```{r}
# Estimate QTE
estimate_QTE <- function(data, tau, obs=TRUE) {
  
  # QTE for experimental study
  if (obs == FALSE) {
    quantile.1 <- estimate_quantile(tau, data[data$treat==1,]$re78, rep(1, nrow(data[data$treat==1,])))
    quantile.0 <- estimate_quantile(tau, data[data$treat==0,]$re78, rep(1, nrow(data[data$treat==0,])))
    
    return(list(qte=quantile.1 - quantile.0, q1=quantile.1, q0=quantile.0))
  }
  
  w.1 <- data$treat / data$ps
  w.0 <- (1 - data$treat) / (1 - data$ps)
  
  quantile.1 <- estimate_quantile(tau, data$re78, w.1)
  quantile.0 <- estimate_quantile(tau, data$re78, w.0)
  
  return(list(qte=quantile.1 - quantile.0, q1=quantile.1, q0=quantile.0))
}
```

```{r}
# Estimate multiple QTE
estimate_QTEs <- function(data, taus, obs = TRUE) {
  # Use sapply to avoid manual indexing
  result <- sapply(taus, function(tau) {
    estimate_QTE(data, tau = tau, obs)$qte
  })
  return(result)
}
```

```{r}
# Test function
taus <- c(0.25, 0.5, 0.75)
estimate_QTEs(re74, taus, FALSE)
```

#### i. For Experimental data

For randomized dataset, QTE equals to the differences between quantiles of the treated and the experimental controls[^4].

[^4]: Firpo, S**.** (2007). Supplement to "Efficient Semiparametric Estimation of Quantile Treatment Effects." *Econometrica*, 75(1), 259–276, p. 261. Supplementary Material, p. 3.

```{r}
# Run functions on experimental datasets and observational datasets before controlling covariates
datasets <- list(re74, psid1_re74, psid2_re74, psid3_re74, cps1_re74, cps2_re74, cps3_re74)

ex_qte_list <- lapply(datasets, function(data) {
  estimate_QTEs(data, taus, FALSE)
})
```

```{r}
# Display result
col_names <- c("QTE 0.25", "0.5", "0.75")
(ex_qte_result <- create_qte_result(ex_qte_list, row_names, col_names))
```

#### **ii. For Observational data**

We use the propensity score estimates from Dehejia and Wahba model, and datasets after discarding some controls.

```{r}
# Run functions using IPW method observational datasets to compute QTE
datasets <- list(filtered_psid1, ate.psid2$trim.data, ate.psid3$trim.data, 
                 filtered_cps1, ate.cps2$trim.data, ate.cps3$trim.data)

qte_results <- lapply(datasets, function(data) {
  estimate_QTEs(data, taus)
})
```

```{r}
col_names <- c("QTE IPW 0.25", "0.5", "0.75")
qte.result_table <- create_qte_result(out_list, row_names, col_names)
```

```{r}
# Display result
cbind(ex_qte_result, qte.result_table)
```

### Summary

-   What did I do?

<!-- -->

-   Why are my results so different from Firpo

<!-- -->

-   What makes QTT and QTE so different?

## Playground

```{r}
# Compute qtes
q <- ci.qte(re78~treat, xformla=~age + I(age^2) + education + I(education^2) +
  married + nodegree + black + hispanic + re74 + I(re74^2) +
  re75 + I(re75^2) + u74 * black , se=FALSE, data=filtered_psid1)
summary(q)
```

```{r}
# Compute qtt
qtt <- ci.qtet(re78~treat, xformla=~age + I(age^2) + education + I(education^2) +
  married + nodegree + black + hispanic + re74 + I(re74^2) +
  re75 + I(re75^2) + u74 * black , se=FALSE, data=psid1_re74)

summary(qtt)
```
