---
title: "R Notebook"
output: html_notebook
---

# Double Score Matching for QTE

Following matching method for QTE and QTT from Yunsun (2021), I directly use their functions to calculate the QTE and QTT for trimmed PSID-1 and CPS-1 with propensity score estimated by the single index model. The prognostic score will be estimated using a GLM model with the same predictors as the single index model.

## QTE

```{r}
taus <- c(0.25, 0.5, 0.75)

# Function to streamline matching process
run_matching_qte <- function(data, covariates, additional_terms, ps_column, taus) {
  Y <- data[,"re78"]
  X <- as.matrix(data[, covariates])
  A <- data[,"treat"]
  Z <- cbind(X, additional_terms(X))  # Add non-linear terms
  sapply(taus, function(tau) {
    dsmatchQTE(Y, X, A, p=tau, method = "dsm", ps=data[[ps_column]], model.pg="linpred", lp.pg=Z)$est.ds
  })
}
```

### Matching for Trimmed PSID-1

```{r}
matching_qte_psid1_SI <- run_matching_qte(
  data = trim_psid1_SI,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"education"]^2, X[,"re74"]^2, X[,"re75"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

### Matching for Trimmed CPS-1

```{r}
matching_qte_cps1_SI <- run_matching_qte(
  data = trim_cps1_SI,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75", "u74", "u75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"age"]^3, X[,"education"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

## QTT

```{r}
taus <- c(0.25, 0.5, 0.75)

# Function to streamline matching process
run_matching_qtt <- function(data, covariates, additional_terms, ps_column, taus) {
  Y <- data[,"re78"]
  X <- as.matrix(data[, covariates])
  A <- data[,"treat"]
  Z <- cbind(X, additional_terms(X))  # Add non-linear terms
  sapply(taus, function(tau) {
    dsmatchQTT(Y, X, A, p=tau, method = "dsm", ps=data[[ps_column]], model.pg="linpred", lp.pg=Z)$est.ds
  })
}
```

### Matching for Trimmed PSID-1

```{r}
matching_qtt_psid1_SI <- run_matching_qtt(
  data = trim_psid1_SI,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"education"]^2, X[,"re74"]^2, X[,"re75"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

### Matching for Trimmed CPS-1

```{r}
matching_qtt_cps1_SI <- run_matching_qtt(
  data = trim_cps1_SI,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75", "u74", "u75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"age"]^3, X[,"education"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

### Matching for Filtered PSID-1

```{r}
matching_qtt_psid1_logi <- run_matching_qtt(
  data = filtered_psid1,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"education"]^2, X[,"re74"]^2, X[,"re75"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

### Matching for Filtered CPS-1

```{r}
matching_qtt_cps1_logi <- run_matching_qtt(
  data = filtered_cps1,
  covariates = c("age", "education", "black", "hispanic", "married", "nodegree", "re74", "re75", "u74", "u75"),
  additional_terms = function(X) {
    cbind(X[,"age"]^2, X[,"age"]^3, X[,"education"]^2)
  },
  ps_column = "ps",
  taus = taus
)
```

## Display result

```{r}
matching_qtt_results_SI <- list(matching_qtt_psid1_SI, matching_qtt_cps1_SI)
matching_qtt_results_logi <- list(matching_qtt_psid1_logi, matching_qtt_cps1_logi)
```

```{r}
col_names <- c("QTT IPW 0.25", "0.5", "0.75")

matching_qtt.result_table.logistic <- create_qte_result(matching_qtt_cps1_logi,  c("NSW", "PSID-1", "CPS-1"), col_names)

matching_qtt.result_table.SI <- create_qte_result(c(matching_qtt_psid1_SI,matching_qtt_psid1_SI),  c("NSW", "PSID-1", "CPS-1"), col_names)

(qtt.result_table <- cbind(ex_qte_result, matching_qtt.result_table.SI, matching_qtt.result_table.logistic))
```
